Analyzing the impact of social attributes on commit integration success

### 2017-02-20 Paper/MSRChallenge2017.tex
getting it to 4 pages

Over the years, the open source
community has developed various techniques to vet project contributions, such as pull
requests which must be approved by a core team member~\cite{gousios14}. Another
contribution mechanism these projects use is continuous
integration, which notifies the team members if a change caused a test 
failure~\cite{Meyer14}.
Yet, these methods can become a bottleneck.
To avoid a possible bottleneck, it would be beneficial to prioritize certain
contributions.  This prioritize can be achieved with predictors of the
contribution's quality.

@@

While past studies have analyzed pull request
success~\cite{tsay14icse,tsay14fse}, we are interested in analyzing the quality 
of a contribution 
at a lower granulairity level: individual commits. \todo{Zack question for Mau:
did we really do the investigation at the individual commit level?} 
We analyzed commits in the 
context of continuous integration, using the technical factors found in the 
some of the social attributes provided in the challenge
corpus~\cite{msr17challenge}, and the social factors found on a developer's
GitHub profile.  We used the technical factors to estimate the complexity of the
commit, and the social factors to estimate the developer's experience.

This study investigates how well different technical and social factors predict
commit quality.  This information will help core team members prioritize project
contributions.  It will also provide a further understand of quality enforcement
in the software development process.

@@

\textbf{Technical Factors:} We collected these factors from the 
MSR Challenge data set~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy the commit's
complexity, such as the size of the changes being performed.  The technical
factors used are listed below:
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item Source churn: lines of code changed by this commit
\item Files added: files added by this commit
\item Files modified: files modified by this commit
\item Tests added: tests added by this commit
\item Tests deleted: tests deleted by this commit
\item Source files: files in project at the time of the commit
\item Source loc: lines of source code in the project at
	the time of the commit.
\item Travis status: status of the Travis build integration
\end{itemize}

@@

We also added one item from the MSR Challenge data set to the social factors:
gh\_by\_core\_team\_member.
This factor indicates if the commit's developer is a core member 
of that project.

@@

The data set used in this study~\cite{msr17challenge} is very diverse.
and consists of commits from multiple years ago. The social data we 
gather is recent, gathered in January 2017. In an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the data set that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
a perfect solution , we believe that these numbers are unlikely to have changed
significantly in the last three years.\clg{...did you spot check? Mau's Response: Spot
check that they did belong to those years? Yeah. Zack's Response: assuming you
are asking a different question of if the values changed in the last three
years, than no, we did not check because we weren't sure how.  What should we
say here?}

\todo{Zack: Mau, check this paragraph}
Since we are interested in the success of the job status of each commit, we 
converted the Travis build statuses to just two possible values,
successful (passed in the original provided data set) and not successful 
(errored, failed, canceled
and started in the original data set). We also condensed down all passing and
failing builds in a commit to a single pass or fail for each build status.  If
the commit contained both passing and failing build statuses, we removed the
commit from the data set.
\todo{confirm that this number doesn't change}
At this point we had a 3:1 successful to 
not successful ratio in the rows of the data sets.

In the non-downsampled data set, 74.43\% of the builds were successful, the rest were not
successful.  This means that a classifier which guesses successful every time
would produce a classification rate of 74.43\% if ran on the full data
set.

The high ratio of successful to not successful builds cause the classifier to 
favor misclassifying not successful builds over successful builds, because a 1\%
misclassification rate in 3/4th of the data is greater than a 1\%
misclassification rate in 1/4th of the data.  To penalize misclassification for
both classes equally, we created a separate data set where the
successful builds equaled the number of not successful builds.  
We build this data set by copying over all not successful builds and randomly
selecting a sub-sample of the successful builds until the two build status were
equal, known as downsampling.  We used 
both the downsampled and non-downsampled data sets in our results to show the 
difference of the behavior of the decision tree when it is trained with a equal 
ratio than when it is trained with a unequal ratio.

\section{Evaluation}

\begin{table*}[ht]
\centering
\begin{tabular}{l||r|r||r|r}
  & \multicolumn{2}{|c||}{Non-Downsampled}  & \multicolumn{2}{|c}{Downsampled} \\
\hline
  & Technical & Technical and Social & Technical & Technical and Social \\
\hline
\textbf{Correctly Classified Instances}&  77.01\% & 81.10\%  & 64.10\% & 76.09\%\\

@@

%\textbf{Kappa statistic}& 0.21 &  0.42 & 0.28  & 0.52 \\

%\textbf{Mean absolute error}& 0.34 &  0.28 & 0.43 & 0.32\\

@@

\hline
\end{tabular}
\\
\center
	\caption{\label{resultsTable}Comparison of predictive power by running 10-fold
    cross validation of two decision trees with a non-downsampled dataset (2nd and 3rd column) and downsampled (4th and 5th column). 
	%The higher the percentages represent the higher number of instances we are able to correctly predict. \label{resultsTable}
	%Zack: the above sentence isn't true for the second row.
	}   
\end{table*}
To give an address to main research question: ``Do social attributes, gathered 
from the GitHub profiles of developers,
allow for more accurate predictions than only using technical data?", we 
created a decision tree classifier~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka J48 decision tree~\cite{Weka,Quinlan1993}
with default settings. The decision trees were evaluated using 10-fold cross 
validation with the data in each data set.

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm, first without downsampling. The 
decision tree built from
only the technical indicators has a correct classification rate of 77.01\% which
means the technical attributes only increased the prediction accuracy by about 3\%
compared to a base line rate of guessing the build passed in all cases.
Adding the social data increased the classifier's accuracy from 77.01\% to
81.10\%, a 4.09\% increase.

On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, which outperformed the baseline of
guessing all builds passed (50.00\%).
 After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data.  These results show that
both technical and social data can be used to improve predictions of build 
integration success.

@@

\paragraph{Internal validity}
The researchers in this study used the dataset provided by the 
 MSRChallenge~\cite{msr17challenge} to gain useful information regarding the 
 behavior of developers when pushing commits and how well these commits integrate 
 successfully with the project. The authors have used 
 Weka,\footnote{http://www.cs.waikato.ac.nz/ml/weka/} the well known tool for 
 usage of machine learning algorithms to analyze our datasets. We are aware that, 
 as every human written tool, Weka is prone to contain errors and/or bugs; 
 nonetheless, Weka is a broadly used, open source tool that is subject to high 
 quality standards and a vast user and developer base which makes it more likely 
 to be a high quality code and to have less bugs than lesser used tools.

 The scripts we have created to adapt and extract the datasets we needed, and the 
 scripts to obtain the developer social data are publicly available to facilitate 
 the transparency and quality of the process(Will be made available after double blind review).%\footnote{https://goo.gl/CqTX44}

\paragraph{External validity} 
 To tackle the concern that the results of this study might not generalize to 
 other instances outside of the studied cases, we have used a broad and rich 
 dataset~\cite{msr17challenge} in our analysis that comprises 1,359 projects. The 
 projects used in this dataset have also been filtered to be popular (having more 
 than 10 watches on GitHub), and to have a history of using the continuous 
 integration tool Travis CI for more than 50 builds. All these features help to 
 build the notion that since we are building our analysis on top of a vast 
 dataset comprehending a wide diversity of projects, the threat of this analysis 
 not generalizing to other projects is mitigated.

\paragraph{Construct validity}
 Travis job success is not the only way to analyze the success of a commit, and 
 that for different reasons the job may not be successful even if the code 
 committed does not have flaws (a flacky test case for example). Since continuous 
 integration is a best known practice in industry and the use of this methodology 
 spreads out into the open source community, we consider that cases of misuse of 
 CI such that it will lead to failing Travis jobs when they should have actually 
 be successful are minimum, specially since having a solid build is one of the 
 main purposes of the methodology, and taking into account that the projects in 
 out dataset are projects with a history of using continuous integration and 
 having a certain level of popularity, this threat is minimized. We therefore 
 consider Travis job success to be a proxy 
 for success when integrating a commit into a larger project.

>>>>

Over the years, the open source
community has developed various techniques to vet project contributions, such as pull
requests which must be approved by a core team member~\cite{gousios14}. 
These methods can become a bottleneck, and to avoid a possible bottleneck, it is beneficial to prioritize certain
contributions.  This prioritization can be achieved with predictors of the
contribution's quality.

@@

While past studies have analyzed pull request
success~\cite{tsay14icse,tsay14fse}, we are interested in analyzing the quality 
of a contribution in the context of continuous integration, we analyze the success every time a developer pushes a contribution to be built.
We analyzed contributions in the 
context of continuous integration, using the technical factors found in
some of the social attributes provided in the challenge
corpus~\cite{msr17challenge}, and the social factors found on a developer's
GitHub profile.  We used the technical factors to estimate the complexity of the
commit, and the social factors to estimate the developer's experience.

This study investigates how well different technical and social factors predict
contribution quality.  This information will help core team members prioritize project
contributions, and will also provide a further understand of quality enforcement
in the software development process.

@@

\textbf{Technical Factors:} We collected these factors from the 
MSR Challenge data set~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy the commit's
complexity, such as the size of the changes being performed.  The technical
factors used are listed below:
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item Source churn: lines of code changed by this commit.
\item Files added: files added by this commit.
\item Files modified: files modified by this commit.
\item Tests added: tests added by this commit.
\item Tests deleted: tests deleted by this commit.
\item Source files: files in project at the time of the commit.
\item Source loc: lines of source code in the project at
	the time of the commit.
\item Travis status: status of the Travis build integration.
\end{itemize}

@@

We also added one item from the MSR Challenge data set to the social factors that indicates if the commit's developer is a core member 
of that project.

@@

The data set used in this study~\cite{msr17challenge} is very diverse.
and consists of commits from multiple years ago. The social data we 
gather is recent, gathered in January 2017. In an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the data set that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
a perfect solution , we believe that these numbers are unlikely to have changed
significantly in the last three years.

Since we are interested in the success of the job status of each contribution, we 
converted the Travis build statuses to just two possible values,
successful (passed in the original provided data set) and not successful 
(errored, failed, canceled
and started in the original data set). We also condensed down all passing and
failing builds in a commit to a single pass or fail for each build status.  If
the commit contained both passing and failing build statuses, this means it is a failed build and we qualify it as such.
At this point 74.43\% of the builds were successful, the rest were not
successful.  This means that a classifier which guesses successful every time
would produce a classification rate of 74.43\% if ran on the full data
set.

The high ratio of successful to not successful builds cause the classifier to 
favor misclassifying not successful builds over successful builds, because a 1\%
misclassification rate in 3/4th of the data is greater than a 1\%
misclassification rate in 1/4th of the data.  To penalize misclassification for
both classes equally, we created a separate data set where the
successful builds equaled the number of not successful builds.  
We build this data set by copying over all not successful builds and randomly
selecting a sub-sample of the successful builds until the two build status were
equal, a process known as downsampling.  We used 
both the downsampled and non-downsampled data sets in our results to show the 
difference of the behavior of the decision tree when it is trained with a equal 
ratio than when it is trained with a unequal ratio.

\section{Evaluation}
\begin{table*}[ht]
\centering
\begin{tabular}{l||r|r||r|r}
  & \multicolumn{2}{|c||}{Non-Downsampled}  & \multicolumn{2}{|c}{Downsampled} \\
\hline
  & Technical & Technical and Social & Technical & Technical and Social \\
\hline
\textbf{Correctly Classified Instances}&  77.01\% & 81.10\%  & 64.10\% & 76.09\%\\

@@

\textbf{Kappa statistic}& 0.21 &  0.42 & 0.28  & 0.52 \\

\textbf{Mean absolute error}& 0.34 &  0.28 & 0.43 & 0.32\\

@@

\hline
\end{tabular}
\\
\center
	\caption{\label{resultsTable}Comparison of predictive power by running 10-fold
    cross validation of two decision trees with a non-downsampled dataset (2nd and 3rd column) and downsampled (4th and 5th column). 
	%The higher the percentages represent the higher number of instances we are able to correctly predict. \label{resultsTable}
	%Zack: the above sentence isn't true for the second row.
	}   
\end{table*}
To address our main research question: ``Do social attributes, gathered 
from the GitHub profiles of developers,
allow for more accurate predictions than only using technical data?", we 
created a decision tree classifier~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka J48 decision tree~\cite{Weka,Quinlan1993}
with default settings. The decision trees were evaluated using 10-fold cross 
validation with the data in each data set.

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm, first without downsampling. The 
decision tree shows an increase of 4.09\% correctly classified instances when adding social data. On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, which outperformed the baseline of
guessing all builds passed (50.00\%).
 After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data.  These results show that
both technical and social data can be used to improve predictions of build 
integration success.

It is also worth noticing that in both cases the Kappa statistic increases when adding social data (100\% and 85\% respectively) which represents a more robust measure than Correctly classified instances, since it takes into account the correctly guessed instances occurring by chance, and we may also notice that the Mean absolute error decreases, giving stronger confidence to our claims when considering eventual outcomes.

@@

%\paragraph{Internal validity}
 The authors used 
 Weka,\footnote{http://www.cs.waikato.ac.nz/ml/weka/} the well known tool for 
 usage of machine learning algorithms to analyze our datasets. We are aware that, 
 as every human written tool, Weka is prone to contain errors and/or bugs; 
 nonetheless, Weka is a broadly used, open source tool that is subject to high 
 quality standards and a vast user and developer base which makes it more likely 
 to be a high quality code and to have less bugs than lesser used tools.

 The scripts we have created to adapt and extract the datasets we needed, and the 
 scripts to obtain the developer social data are publicly available to facilitate 
 the transparency and quality of the process (Will be available after double blind review).%\footnote{https://goo.gl/CqTX44}

%\paragraph{External validity} 
 To tackle the concern that the results of this study might not generalize to 
 other instances outside of the studied cases, we have used a broad and rich 
 dataset~\cite{msr17challenge} in our analysis that comprises 1,359 popular projects, which help to 
 build the notion that since we are building our analysis on top of a vast 
 data set, the threat of this analysis 
 not generalizing to other projects is mitigated.
